---
title: "Project PA"
output: html_document
date: "2023-01-12"
---

```{r}
library("gbm")
library("dplyr")
library("gratia")
library("ggeffects")
library("Hmisc")
library("caret")
library("randomForest")
library("rpart")
library("ggplot2")
library("tidyverse")
library("gam")
```

```{r}
train=read.csv("D:/train.csv",sep=";",header=T)
train=select(train,-torque)
train
str(train)
```

#train name variable
```{r}
train$name <- sapply(strsplit(train$name, " "), `[`, 1)
ggplot(data = train, aes(x=name, fill = name)) +
  geom_bar() + labs(x='train Brand') + labs(title = "Bar Graph of train Brand") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
train$name <- str_replace(train$name, 'Maruti', '0')
train$name <- str_replace(train$name, 'Skoda', '1')
train$name <- str_replace(train$name, 'Honda', '2')
train$name <- str_replace(train$name, 'Hyundai', '3')
train$name <- str_replace(train$name, 'Toyota', '4')
train$name <- str_replace(train$name, 'Ford', '5')
train$name <- str_replace(train$name, 'Renault', '6')
train$name <- str_replace(train$name, 'Mahindra', '7')
train$name <- str_replace(train$name, 'Tata', '8')
train$name <- str_replace(train$name, 'Chevrolet', '9')
train$name <- str_replace(train$name, 'Fiat', '10')
train$name <- str_replace(train$name, 'Datsun', '11')
train$name <- str_replace(train$name, 'Jeep', '12')
train$name <- str_replace(train$name, 'Mercedes-Benz', '13')
train$name <- str_replace(train$name, 'Mitsubishi', '14')
train$name <- str_replace(train$name, 'Audi', '15')
train$name <- str_replace(train$name, 'Volkswagen', '16')
train$name <- str_replace(train$name, 'BMW', '17')
train$name <- str_replace(train$name, 'Nissan', '18')
train$name <- str_replace(train$name, 'Lexus', '19')
train$name <- str_replace(train$name, 'Jaguar', '20')
train$name <- str_replace(train$name, 'Land', '21')
train$name <- str_replace(train$name, 'MG', '22')
train$name <- str_replace(train$name, 'Volvo', '23')
train$name <- str_replace(train$name, 'Daewoo', '24')
train$name <- str_replace(train$name, 'Kia', '25')
train$name <- str_replace(train$name, 'Force', '26')
train$name <- str_replace(train$name, 'Ambassador', '27')
train$name <- str_replace(train$name, 'Ashok', '28')
train$name <- str_replace(train$name, 'Isuzu', '29')
train$name <- str_replace(train$name, 'Opel', '30')
train$name <- str_replace(train$name, 'Peugeot', '31')
```

```{r}
train$name <- as.numeric(train$name)
table(train$name)
```
#Substituting blank with NA for columns mileage, engine, max_power
```{r}
train$mileage[train$mileage == ""] <- NA
train$engine[train$engine == ""] <- NA
train$max_power[train$max_power == ""] <- NA
```

#Checking for missing values
```{r}
sapply(train, function(x) sum(is.na(x)))
```
```{r}
#Removing unit from mileage, converting it to numeric value and replacing the missing values
train$mileage <- str_replace(train$mileage, 'kmpl', '')
train$mileage <- str_replace(train$mileage, 'km/kg', '')
train$mileage <- as.numeric(train$mileage)
train$mileage[is.na(train$mileage)]<-mean(train$mileage,na.rm=TRUE)

#Removing unit from engine, converting it to numeric value and replacing the missing values
train$engine <- str_replace(train$engine, 'CC', '')
train$engine <- as.numeric(train$engine)
train$engine[is.na(train$engine)]<-mean(train$engine,na.rm=TRUE)

#Removing unit from max_power, converting it to numeric value and replacing the missing values
train$max_power <- str_replace(train$max_power, 'bhp', '')
train$max_power <- as.numeric(train$max_power)
train$max_power[is.na(train$max_power)]<-mean(train$max_power,na.rm=TRUE)

#Converting seats to numeric value and replacing the missing values
train$seats <- as.numeric(train$seats)
train$seats[is.na(train$seats)]<-median(train$seats,na.rm=TRUE)
```

```{r}
# Checking for missing values once again
sapply(train, function(x) sum(is.na(x)))
```

#Plotting categorical Values and checking for distribution
```{r}
# Bar graph of Fuel
ggplot(data = train, aes(x=reorder(fuel, fuel, function(x)-length(x)), fill = fuel)) +
  geom_bar() + labs(x='Fuel') + labs(title = "Bar Graph of Fuel") 
```
```{r}
#Bar graph of Seller Typs
ggplot(data = train, aes(x=reorder(seller_type, seller_type, function(x)-length(x)), fill = seller_type)) +
  geom_bar() + labs(x='Seller Type') + labs(title = "Bar Graph of Seller Type")
```
Huge number of cars are owned by individual owners followed by Dealer and Trustmark Dealers.

```{r}
# Bar graph of Owner
ggplot(data = train, aes(x=reorder(owner, owner, function(x)-length(x)), fill = owner)) +
  geom_bar() + labs(x='Owner') + labs(title = "Bar Graph of Owner") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Most of the cars are owned by first owners.

# Bar graph of seats
```{r}
ggplot(data = train, aes(x=reorder(seats, seats, function(x)-length(x)), fill = seats)) +
  geom_bar() + labs(x='Seats') + labs(title = "Bar Graph of Seats") +theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```
Most of the cars are 5 seater

#Converting transmission, owner, seller type and fuel into ordinal encoder
```{r}
#Converting transmission column into binary 0 if Manual and 1 if Automatic
train$transmission <- str_replace(train$transmission, 'Manual', "0")
train$transmission <- str_replace(train$transmission, 'Automatic', "1")
train$transmission <- as.numeric(train$transmission)
table(train$transmission)
```

#Converting owner into Ordinal Encoder
```{r}
train$owner <- str_replace(train$owner, 'First Owner', "0")
train$owner <- str_replace(train$owner, 'Second Owner', "1")
train$owner <- str_replace(train$owner, 'Third Owner', "2")
train$owner <- str_replace(train$owner, 'Fourth & Above Owner', "3")
train$owner <- str_replace(train$owner, 'Test Drive Car', "4")
train$owner <- as.numeric(train$owner)
table(train$owner)
```
#Converting seller_type into Ordinal Encoder
```{r}
train$seller_type <- str_replace(train$seller_type, "Trustmark Dealer", "0")
train$seller_type <- str_replace(train$seller_type, "Dealer", "1")
train$seller_type <- str_replace(train$seller_type, "Individual", "2")
train$seller_type <- as.numeric(train$seller_type)
table(train$seller_type)
```
#Converting fuel into Ordinal Encoder
```{r}
train$fuel <- str_replace(train$fuel, 'Diesel', "0")
train$fuel <- str_replace(train$fuel, 'Petrol', "1")
train$fuel <- str_replace(train$fuel, 'CNG', "2")
train$fuel <- str_replace(train$fuel, 'LPG', "3")
train$fuel <- as.numeric(train$fuel)
table(train$fuel)
```
#Plotting histogram of selling price, km driven to check the distribution
```{r}
#Histogram of Selling Price
ggplot(train, aes(x=selling_price)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="blue")+
  labs(x='Selling Price ') + labs(title = "Histogram Graph of Selling Price") +
  scale_x_continuous(trans='log10')
```

```{r}
#Histogram of Km Driven
ggplot(train, aes(x=km_driven)) + 
  geom_histogram(color="black", fill="blue", bins = 200)+
  labs(x='Km Driven ') + labs(title = "Histogram Graph of Km Driven") +
  scale_x_continuous(trans='log10')
```

```{r}
res2 <- rcorr(as.matrix(train))
res2
```
We can see that selling price is highly correlated to max_power then transmission and name.


#Splitting the Data into training and test data sets
```{r}
set.seed(5)
trainIndex <- createDataPartition(train$selling_price, p = .8,
                                  list = FALSE,
                                  times = 1)
Train <- train[ trainIndex,]
Test <- train[-trainIndex,]
```
Splitting data into 80% Training and 20% Test.    

#GAM
```{r}
gam1 <- gam(selling_price ~ s(name) + s(year) + s(km_driven) + seller_type + transmission  + s(mileage) + s(engine) + s(max_power) , data = Train)
summary(gam1)

predict_gam <- predict(gam1, Test)
error_gam <- Test$selling_price - predict_gam
RMSE_gam <- sqrt(mean(error_gam^2))
RMSE_gam
```

#RANDOM FOREST
```{r}
rf=randomForest(selling_price~.,data = Train)
rf

#Using the model to predict selling price in the Test dataset
pred_rf <- predict(rf, Test)
error_rf <- Test$selling_price - pred_rf
RMSE_rf <- sqrt(mean(error_rf^2))
RMSE_rf

#Plotting predicted vs. actual values
plot(Test$selling_price,pred_rf, main="Scatterplot", col = c("red","blue"), xlab = "Actual Selling Price", ylab = "Predicted Selling Price")

varImpPlot(rf, main ='Feature Importance')
```


#GRADIENT BOOSTING
```{r}
set.seed(123)
gbm <- gbm(
  formula = selling_price ~ .,
  distribution = "gaussian",
  data = Train,
  n.trees = 6000,
  interaction.depth = 3,
  shrinkage = 0.1,
  cv.folds = 5,
  n.cores = NULL, 
  verbose = FALSE
)  
gbm
```
```{r}
#Using the model to predict selling price in the Test dataset
pred_gbm <- predict(gbm, Test)
error_gbm <- Test$selling_price - pred_gbm
RMSE_gbm <- sqrt(mean(error_gbm^2))
RMSE_gbm

summary(gbm, cBars = 10,method = relative.influence, las = 2)
```

```{r}
cp=data.frame("Model"= c("GAM","RANDOM FOREST","BOOSTING"),
              "RMSE"=c(RMSE_gam,RMSE_rf,RMSE_gbm))
cp
```
We used Generalized additive model, random forest and gradient boosting models to predict selling price of cars and we see that gradient boosting gives us a better RMSE among the three models.Variables that are useful to describe the variance are max_power, year, km_driven and name.





